---
title: "p8105_hw3_nl2835"
author: "Nancy Le"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```


```{r load_libraries}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
data("instacart")
```

### Problem 1

The Instacart dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables, which include `order_id`, `product_id`, the `user_id`, `product_name`s, and which location of the supermarket the products were found (as in, the `department` and `aisle`). 

User IDs ranged from 1 to 206209. 

```{r count aisles}
num_aisles <- instacart |> 
  select(aisle) |> 
  distinct() |> 
  nrow()

print(num_aisles)
```

There are 134 aisles. 


```{r most items}
count_items_aisle <- instacart |> 
 group_by(aisle_id) |> 
 count() |> 
  arrange(desc(n))
```


The aisle with the most items ordered from is `r count_items_aisle$aisle_id[1]`.


```{r collapse instacart df}
instacart = instacart |> 
  group_by(aisle_id) |> 
  mutate(total_product_count = n()) |> 
  select(order_id, aisle_id, product_id, total_product_count, everything())
```

```{r filter instacart data}
subset_instacart = instacart |> 
  filter(total_product_count > 10000) 
```

```{r scatterplot items by aisle}
subset_instacart |> 
  ggplot(aes(x = aisle_id, y = total_product_count)) + geom_point() + labs(
    title = "Number of Products Ordered by Aisle",
    x = "Aisle ID",
    y = "Number of Products Ordered",
    caption = "Data from Instacart") +
  scale_x_continuous(
    breaks = seq(0, 135, by = 1)
  )

#need to go back and fix aisle id labels - maybe should use text labels above just the top 3 product counts?
```


```{r bar plot items by aisle}
subset_instacart |> 
ggplot(aes(x = aisle_id, y = total_product_count)) + 
  geom_bar(stat = "identity") + 
  scale_x_discrete(labels = subset_instacart$aisle_id)
```

```{r table}
three_depts = instacart |> 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(product_id, aisle_id) 

```

```{r}
three_depts = three_depts |> 
  mutate(product_count = 1) |> 
  select(aisle_id, aisle, product_id, product_name, product_count, everything()) 
```


```{r}
top_three = three_depts |> 
  group_by(aisle_id, product_name) |> 
  summarize(product_count = sum(product_count)) |> 
  arrange(desc(product_count))  |> 
  slice_head(n = 3) |> 
  select(product_name, product_count, aisle_id) 
```


```{r}
library(knitr)
kable(top_three, caption = "Top 3 Products By Aisle")
```

```{r}
days_df <- 
  tibble(
    day_num = 0:6,
    day_abb = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat")
  )
```

```{r}
replacement_vector <- setNames(days_df$day_abb, days_df$day_num)
```

```{r}
pla_cic = instacart |> 
  filter(product_name == c(
    "Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(order_dow, product_name) |>
  summarize(mean_hr = mean(order_hour_of_day)) |> 
  spread(key = order_dow, value = mean_hr) 
```


```{r}
library(knitr)
kable(pla_cic, caption = "Mean Hour Ordered")
```


## Problem 2

```{r brfss load}
data("brfss_smart2010")
```

```{r clean data}
brfss = brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic=="Overall Health") |> 
  drop_na(response) |> 
  mutate(response = as_factor(response)) |> 
  mutate(locationdesc = substring(locationdesc, 5)) |> 
  rename(county = locationdesc, state = locationabbr)
```
